title: "Data Analysis Project"
author: "Marijke Friedman, Brendan Weissel, Sade Ajishegiri"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("ggplot2")
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggthemes)
library(tidycensus)
library(ggplot2)
```
There are a lot of people who were shocked and confused by the results of the election. Over and over again, they reference the dramatic shift of so many states from blue to red, trying to make sense of a country they no longer recognize. One of the keys to understanding what happened, why people voted the way they did, is to investigate counties divided in themselves; the smaller the county, the more significant the split. Kent County might be such a county split between Democratic and Republican voters —the same number of registered Democrats and Republicans voted early, indicating a similar split among those who voted on election day. Here's how we found that:
### [MOVE CODE FROM Q 4 HERE (?)]
Kent County is mostly white — over 81%. The median household income is lower than the national average at $74,402 — especially low for Maryland, whose median average is nearly $99,000. At 12.1%, Kent County’s poverty rate is closer to the national (12.5%) than to the state’s (9.5%).

```{r}
ev_raw <- read_csv("data/EV_RAW_DATA.csv") 

ev_raw <- ev_raw |>
  clean_names()
glimpse(ev_raw)
```

```{r}
#votes by day 
votes_by_day <- ev_raw |> 
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

#votes by party
votes_by_party <- ev_raw |>
  group_by(party_code) |> 
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE)) |> 
  arrange(party_code)

### votes by precinct 
votes_by_precinct <- ev_raw|>
  group_by(precinct)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(precinct)

### votes by gender
votes_by_gender <- ev_raw|>
  group_by(gender_code)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(gender_code)

### votes by age
votes_by_age <- ev_raw|>
  group_by(age_group)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(age_group)
```

### Our original questions:

1. What age group had the highest turn out for early voting?
2. What county had the highest turnout for early voting?
3. How does the 2024 early voting data compare to the 2020 early voting data?
4. What counties had the closest margin between registered Republicans and Democrats who voted early? Can we identify any swing counties from this?
5. How do early voting patterns vary among different demographic groups? 

###: Our new questions:

1. What precincts did younger age groups dominate voting early in? What do these precincts have in common?
2. Are there any precincts where certain demographic groups outperformed their Census data? For example, are there precincts were 65% of the early voters were women despite women only being 50% of the voting age population?
3. Which precincts had the most early voters identifying outside of male/female? How did they vote?
4. Which precincts have historically turned out more early voters than others? How did it compare this election?
5. Do most precincts follow the general pattern of having the most people vote on the first and last days of early voting? 

### How we've refined our questions:

Based on your feedback, we're refining our questions to focus more on precinct-level data instead of overall state data. We are looking more closely at patterns that stand out in certain precincts - taking note of the unusual and trying to find the reasons behind it.

### Describing any obvious limitations/flaws in our data - Why might we be unable to answer our previous questions? Describe the limits of the data in terms of the time frame, what is and is not included and any codes or other information we'll need to find definitions for, and what we've done to better define it. 

The data we have is just early voting data. It's interesting to analyze because early voters are likely to be more civic duty-minded. But the early voter data doesn't tell us that much about the overall electorate, so it could be helpful to bring in actual results data and more information about the overall electorate.

We'll likely use Census data so we can make more meaningful comparisons, plus voter registration data so we can look at voter turnout numbers. 

As far as defining some of the terms in our questions and within our data, in the votes_by_gender table, what does "NA" actually mean? Did voters in that category mark a box? Did they write in something other than male or female? Was there an option that said "prefer not to say"?

### Question 1: What voting location and or county did younger age groups dominate voting early in? What do these places have in common?

The state youth turnout % for early voting was 6.55%. 8 Counties had a higher %. 1. Saint Mary's (7.6%), 2. Howard (7.5%), 3. Charles (7.1%), 4. Calvert (7.1%), 5. Montgomery(6.9%), 6. Frederick(6.9%) 7. Prince George's (6.8%), 8. Washington (6.8%). The top counties tend to have a lower median age compared to other counties.

The top polling location for early youth vote is likely due to the proximity to the University of Maryland. 

The young vote percentage in PG County was 6.8% so the College Park Community Center out performed the county average by about 3 points with 10% being youth vote of the locations voting total.

The counties with highest percent of young voters (aged 18-24) did not specifically correlate to early voting youth turnout. For example Wicomico, Somerset, Allegany and Kent county all have more than 10% of their total population in the 18-24 range. However none of them ranked in the top 8 for youth voter turnout. The bar chart below shows how all counties youth population compares to early youth turnout. 

```{r}
#total votes by county
total_votes_by_county <- ev_raw |>
  group_by(county_name) |>
  summarise(
    total_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
  )

#total votes by ages 18-24 by county
young_votes_by_county <- ev_raw |>
  filter(age_group == "18-24") |>
  group_by(county_name) |>
  summarise(
    young_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
  )

#calculating the voter share or percentage that young voters have in each county
young_voters_by_county_by_percent <- young_votes_by_county |>
  left_join(total_votes_by_county, by = "county_name") |>
  mutate(young_vote_share = young_votes / total_votes) |>
  arrange(desc(young_vote_share))

#total votes by early voting location
total_votes_by_location <- ev_raw |>
  group_by(early_vote_center, county_name) |>
  summarise(
    total_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  )

#total votes by ages 18-24 by early voting location
young_votes_by_location <- ev_raw |>
  filter(age_group == "18-24") |>
  group_by(early_vote_center) |>
  summarise(
    young_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  )

young_voters_by_location_by_percent <- young_votes_by_location |>
  left_join(total_votes_by_location, by = "early_vote_center") |>
  mutate(young_vote_share = young_votes / total_votes) |>
  arrange(desc(young_vote_share))

#Maryland total early youth vote %
total_young_votes <- ev_raw |>
  filter(age_group == "18-24") |>
  summarise(total_young_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE))

total_votes <- ev_raw |>
  summarise(total_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE))

state_youth_turnout <- total_young_votes$total_young_votes / total_votes$total_votes * 100

state_youth_turnout


#18-24 on the census data
age_18_24 <- get_acs(
  geography = "county",
  variables = c(
    "B01001_007", # M 18-19
    "B01001_008", # M 20
    "B01001_009", # M 21
    "B01001_010", # M 22-24
    "B01001_031", # F 18-19
    "B01001_032", # F 20
    "B01001_033", # F 21
    "B01001_034"  # F 22-24
  ),
  state = "MD",
  year = 2022
) |>
  group_by(GEOID) |>
  summarise(youth_pop = sum(estimate))

#total population
total_pop <- get_acs(
  geography = "county",
  variables = "B01001_001", 
  state = "MD",
  year = 2022
)

#calculate percentage
youth_percent <- total_pop |>
  left_join(age_18_24, by = "GEOID") |>
  mutate(
    NAME = gsub(", Maryland", "", NAME),
    NAME = gsub(" County", "", NAME),
    NAME = case_when(
      NAME == "St. Mary's" ~ "Saint Mary's",
      NAME == "Baltimore" ~ "Baltimore County",
      NAME == "Baltimore city" ~ "Baltimore City",
      TRUE ~ NAME
    ),
    youth_percent = (youth_pop / estimate) * 100
  ) |>
  select(
    county_name = NAME,
    total_pop = estimate,
    youth_pop,
    youth_percent
  ) |>
  arrange(desc(youth_percent))

#combine census data with early voting data
Youth_voting_census_comparison <- young_voters_by_county_by_percent |>
  mutate(
    county_name = case_when(
      county_name == "St. Mary's" ~ "Saint Mary's",
      county_name == "Baltimore" ~ "Baltimore County",
      county_name == "Baltimore city" ~ "Baltimore City",
      TRUE ~ county_name
    )
  ) |>
  left_join(youth_percent, by = "county_name") |>
  mutate(
    young_vote_share = young_vote_share * 100,
  ) |>
  select(
    county_name,
    young_vote_share,
    census_youth_percent = youth_percent,
    total_votes,
    young_votes,
    census_youth_pop = youth_pop,
    census_total_pop = total_pop
  ) |>
  arrange(desc(census_youth_percent))


youth_comparison_clean <- Youth_voting_census_comparison |>
 select(county_name, young_vote_share, census_youth_percent) |>
 rename(
   "Youth Turnout" = young_vote_share,
   "Youth Population" = census_youth_percent
 ) |>
 pivot_longer(
   cols = c("Youth Turnout", "Youth Population"),
   names_to = "metric",
   values_to = "percentage"
 )

#chart :)
ggplot(youth_comparison_clean) +
 geom_col(aes(x = reorder(county_name, -percentage * (metric == "Youth Population")), 
              y = percentage,
              fill = metric), 
          position = position_dodge()) +
 scale_fill_manual(values = c("Youth Turnout" = "Blue", "Youth Population" = "green")) +
 theme_minimal() +
 labs(
   title = "High Youth Population ≠ High Early Youth Turnout",
   x = "County",
   y = "Percentage",
   fill = "Measure",
   caption = "Source: Maryland State Board of Elections and Census"
 ) +
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Answering Question 2: Are there any precincts where certain demographic groups outperformed their Census data? For example, are there precincts were 65% of the early voters were women despite women only being 50% of the voting age population?

Looking at the overall population split between male and female in each county and the split between male and female in early voting we concluded that Women are over represented in early voting compared to their population share across ALL counties. The largest gaps in female population to were in Somerset (9.3% higher), Prince George's (9.1% higher), and Baltimore City (8.1% higher). The smallest gaps (Worcester at 1.3%, Dorchester at 1.8%) still showed women voting above their population share. 

```{r}
# gender breakdown of early voters by county
ev_gender_breakdown <- ev_raw |>
 mutate(
   county_name = case_when(
     county_name == "St. Mary's" ~ "Saint Mary's",
     county_name == "Baltimore" ~ "Baltimore County",
     county_name == "Baltimore city" ~ "Baltimore City",
     TRUE ~ county_name
   )
 ) |>
   filter(gender_code != "N/A") |>  
 group_by(county_name, gender_code) |>
 summarise(
   votes_by_gender = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 group_by(county_name) |>
 mutate(
   total_votes = sum(votes_by_gender),
   voting_gender_percent = round((votes_by_gender/total_votes) * 100, 3)
 ) |>
 arrange(desc(voting_gender_percent))

# Census gender population data 
gender_population_data <- get_acs(
 geography = "county",
 variables = c(
   male = "B01001_002",    # Male total
   female = "B01001_026"   # Female total
 ),
 state = "MD",
 year = 2022
)

# Calculate gender percentages using census data
census_gender_percents <- gender_population_data |>
 mutate(
   NAME = gsub(", Maryland", "", NAME),
   NAME = gsub(" County", "", NAME),
   NAME = case_when(
     NAME == "St. Mary's" ~ "Saint Mary's",
     NAME == "Baltimore" ~ "Baltimore County",
     NAME == "Baltimore city" ~ "Baltimore City",
     TRUE ~ NAME
   )
 ) |>
 select(
   county_name = NAME,
   gender_code = variable,
   population = estimate
 ) |>
 group_by(county_name) |>
 mutate(
   total_pop = sum(population),
   census_gender_percent = round((population/total_pop) * 100, 3),
   gender_code = case_when(
     gender_code == "male" ~ "M",
     gender_code == "female" ~ "F"
   )
 ) |>
 arrange(county_name, gender_code)

# Compare census gender breakdown by county to early voting gender breakdown by county
gender_comparison <- ev_gender_breakdown |>
 left_join(census_gender_percents, by = c("county_name", "gender_code")) |>
 mutate(
   representation_difference = voting_gender_percent - census_gender_percent
 ) |>
 select(
   county_name,
   gender_code,
   voting_gender_percent,
   census_gender_percent,
   representation_difference,
   votes_by_gender,
   total_votes,
   population,
   total_pop
 ) |>
 arrange(desc(representation_difference))

```

### Answering Question 3: Which precincts had the most early voters identifying outside of male/female? How did they vote?

Non-binary of N/A voters according to the early voting records represent  0.301% of all early voters. Baltimore county (453), Anne Arundel (418), Montgomery (399) and Baltimore City (324) were the counties with the highest number of non-binary voters however when compared to the overall population of the county. However Queen Anne's (8.15 per 10k), Talbot (7.97 per 10k), and Kent (7.78 per 10k) had he most non-binary voters per capita. 

11 Counties had over 0.301% of their voters identify as non-binary. Baltimore City (0.0605%), Kent (0.414%) and Anne Arundel (0.408%) had the highest totals comapred to their overall early vote totals. Also, Democrats had the highest percent of non-binary voters 1629 of the total 2989. 


```{r}
#male/female/n.a totals
gender_vote_totals <- ev_raw |>
 group_by(gender_code) |>
 summarise(
   votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_votes = sum(votes),
   percent_of_votes = round((votes/total_votes) * 100, 3)
 ) |>
 arrange(desc(votes))

# Find precincts with most non binary voters
nonbinary_precinct_analysis <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(precinct, county_name, early_vote_center) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 # Get total voters at each precinct for percentage
 left_join(
   ev_raw |>
     group_by(precinct) |>
     summarise(
       total_precinct_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
     ),
   by = "precinct"
 ) |>
 mutate(
   na_percent = round((na_voters/total_precinct_voters) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# Party breakdown for non-binary early voters
na_party_breakdown <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(party_code) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_na = sum(na_voters),
   party_percent = round((na_voters/total_na) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# County breakdown for non-binary early voters
na_county_breakdown <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(county_name) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_na = sum(na_voters),
   county_percent = round((na_voters/total_na) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# Using population data by county to see what county had the highest percent of non binary voters in relation to total county population
county_pop <- get_acs(
  geography = "county",
  variables = "B01001_001", 
  state = "MD",
  year = 2022
) |>
  mutate(
    NAME = gsub(", Maryland", "", NAME),
    NAME = gsub(" County", "", NAME),
    NAME = case_when(
      NAME == "St. Mary's" ~ "Saint Mary's",
      NAME == "Baltimore" ~ "Baltimore County",
      NAME == "Baltimore city" ~ "Baltimore City",
      TRUE ~ NAME
    )
  )

# Look at county breakdown of non binary voters with census context (using per 10k)
na_county_census_breakdown <- ev_raw |>
  filter(is.na(gender_code)) |>
  group_by(county_name) |>
  summarise(
    na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(county_pop |> select(county_name = NAME, total_pop = estimate), 
            by = "county_name") |>
  mutate(
    total_na = sum(na_voters),
    county_percent = round((na_voters/total_na) * 100, 3),
    na_per_10k = round((na_voters/total_pop) * 10000, 3) 
  ) |>
  arrange(desc(na_per_10k))

# Compare non-binary voters to total voters by county - (In what counties did non-binary voters make up the highest percents)
na_voting_percentages <- total_votes_by_county |>
 left_join(na_county_breakdown, by = "county_name") |>
 mutate(
   percent_na_voters = round((na_voters/total_votes) * 100, 3)
 ) |>
 select(
   county_name,
   total_votes,
   na_voters,
   percent_na_voters
 ) |>
 arrange(desc(percent_na_voters))


```

### Answering Question 4: What counties had the closest margin between registered Republicans and Democrats who voted early? Can we identify any swing counties from this?

We identified two swing counties: Kent County is the most competitive from early results with (42.7% DEM, 42.1% REP, margin of 0.6%). Anne Arundel shows close competition (40.8% DEM, 39.1% REP, margin of 1.7%). 

89.1% of precincts in Maryland were not competitive with a greater than 10% margin between democrats and republicans. 5.5% of precincts could be competitive with a 5-10% margin between democrats and republicans. 4.1% of precincts were very competitive with a 1-5% margin between democrats and republicans. 1.3% or 25 total precincts were extremely competitive with less than 1% margin between democrats and republican. 3 of these precincts only recorded 1 total and neither were for the democratic or republican candidate and 3 other counties recorded 15 or fewer votes. In total we found 19 total very competitive precincts that could be used to predict election results. Baltimore County accounted for 8 of the 19 total very competitive precincts; however the overall county favored Dem by 24 points in the early results. 


```{r}
party_by_county <- ev_raw |>
 group_by(county_name, party_code) |>
 summarise(
   votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 group_by(county_name) |>
 mutate(
   total_votes = sum(votes),
   party_share = round((votes / total_votes) * 100, 1)
 ) |>
 summarise(
   total_votes = first(total_votes),
   votes_REP = sum(votes[party_code == "REP"], na.rm = TRUE),
   votes_DEM = sum(votes[party_code == "DEM"], na.rm = TRUE),
   votes_UNA = sum(votes[party_code == "UNA"], na.rm = TRUE),
   votes_OTH = sum(votes[party_code == "OTH"], na.rm = TRUE),
   votes_LIB = sum(votes[party_code == "LIB"], na.rm = TRUE),
   votes_GRN = sum(votes[party_code == "GRN"], na.rm = TRUE),
   votes_NLM = sum(votes[party_code == "NLM"], na.rm = TRUE),
   party_share_REP = sum(party_share[party_code == "REP"], na.rm = TRUE),
   party_share_DEM = sum(party_share[party_code == "DEM"], na.rm = TRUE),
   party_share_UNA = sum(party_share[party_code == "UNA"], na.rm = TRUE),
   party_share_OTH = sum(party_share[party_code == "OTH"], na.rm = TRUE),
   party_share_LIB = sum(party_share[party_code == "LIB"], na.rm = TRUE),
   party_share_GRN = sum(party_share[party_code == "GRN"], na.rm = TRUE),
   party_share_NLM = sum(party_share[party_code == "NLM"], na.rm = TRUE)
 ) |>
 mutate(
   margin_DEM_REP = abs(party_share_DEM - party_share_REP),
   competitiveness = case_when(
     margin_DEM_REP <= 1 ~ "Extremely Competitive (<1%)",
     margin_DEM_REP <= 5 ~ "Very Competitive (1-5%)",
     margin_DEM_REP <= 10 ~ "Could be Competitive (5-10%)",
     TRUE ~ "Non Competitive (>10%)"
   )
 ) |>
 arrange(margin_DEM_REP)

#now by precinct 
party_by_precinct <- ev_raw |>
 group_by(precinct, county_name, party_code) |>
 summarise(
   votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 group_by(precinct, county_name) |>
 mutate(
   total_votes = sum(votes),
   party_share = round((votes / total_votes) * 100, 1)
 ) |>
 summarise(
   total_votes = first(total_votes),
   votes_REP = sum(votes[party_code == "REP"], na.rm = TRUE),
   votes_DEM = sum(votes[party_code == "DEM"], na.rm = TRUE),
   votes_UNA = sum(votes[party_code == "UNA"], na.rm = TRUE),
   votes_OTH = sum(votes[party_code == "OTH"], na.rm = TRUE),
   votes_LIB = sum(votes[party_code == "LIB"], na.rm = TRUE),
   votes_GRN = sum(votes[party_code == "GRN"], na.rm = TRUE),
   votes_NLM = sum(votes[party_code == "NLM"], na.rm = TRUE),
   party_share_REP = sum(party_share[party_code == "REP"], na.rm = TRUE),
   party_share_DEM = sum(party_share[party_code == "DEM"], na.rm = TRUE),
   party_share_UNA = sum(party_share[party_code == "UNA"], na.rm = TRUE),
   party_share_OTH = sum(party_share[party_code == "OTH"], na.rm = TRUE),
   party_share_LIB = sum(party_share[party_code == "LIB"], na.rm = TRUE),
   party_share_GRN = sum(party_share[party_code == "GRN"], na.rm = TRUE),
   party_share_NLM = sum(party_share[party_code == "NLM"], na.rm = TRUE),
   county_name = first(county_name)
 ) |>
 mutate(
   margin_DEM_REP = abs(party_share_DEM - party_share_REP),
   competitiveness = case_when(
     margin_DEM_REP <= 1 ~ "Extremely Competitive (<1%)",
     margin_DEM_REP <= 5 ~ "Very Competitive (1-5%)",
     margin_DEM_REP <= 10 ~ "Could be Competitive (5-10%)",
     TRUE ~ "Non Competitive (>10%)"
   )
 ) |>
 arrange(margin_DEM_REP)

competitiveness_summary <- party_by_precinct |>
 group_by(competitiveness) |>
 summarise(
   number_of_precincts = n(),
   .groups = "drop"
 ) |>
 mutate(
   percent_of_precincts = round((number_of_precincts/sum(number_of_precincts)) * 100, 1)
 ) |>
 arrange(factor(competitiveness, levels = c(
   "Extremely Competitive (<1%)",
   "Very Competitive (1-5%)",
   "Could be Competitive (5-10%)",
   "Non Competitive (>10%)"
 )))

```


### Answering Question 5: Do most precincts follow the general pattern of having the most people vote on the first AND last days of early voting?  Do most early voting locations follow the general pattern of having the most people vote on the first AND last days of early voting? Do most counties follow the general pattern of having the most people vote on the first AND last days of early voting?

No, most precincts do not follow the pattern of having the most people vote on the first AND last days of early voting. Only about 27% of precincts follow this pattern. For the second question, no most counties do not follow this pattern. The percentage is slightly higher at about 33%. One thing that We did notice through looking at this is that Prince George's County had significantly higher voters on the last day of early voting. Statewide, the first and last days of early voting had the most votes cast. If we look at this by polling locations, we can see that 36% of the early voting locations follow the pattern. We can also see that 14 of 24 counties have at least one polling location that follows the pattern. Among them are some of the counties with the most polling locations (Anne Arundel, Prince George's, Montgomery, Baltimore County, Baltimore City). These are some of the more populous regions in the state and some of those locations have more voters at them than others. This could be why statewide the first and last days of early voting had the most voters in Maryland, but this pattern was not followed when broken down by precinct, county, and polling location. 

```{r}

### BY POLLING LOCATION

location_by_day_simple <- ev_raw|>
  group_by(early_vote_center, county_name) |>
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

location_by_day <- location_by_day_simple |>
  rowwise() |>
  mutate(
    first_day = `day1`,             
    last_day = `day8`,           
    max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[1],  
    second_max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[2],  
    follows_pattern = first_day %in% c(max_votes, second_max_votes) &
                      last_day %in% c(max_votes, second_max_votes) &
                      first_day != last_day)

summing_locations <- location_by_day |>
  mutate(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))

location_percentage_follows_pattern <- location_by_day |>
  ungroup() |> 
  summarize(
    location_following = sum(follows_pattern),
    total_location = n(),
    percentage = (location_following / total_location) * 100
  )

### creating line chart for polling locations by day

location_by_day_long <- location_by_day_simple |>
  pivot_longer(
    cols = starts_with("day"),
    names_to = "day",
    values_to = "votes"
  ) |>
  mutate(day = as.numeric(gsub("day", "", day)))  # Extract numeric day

print(
  ggplot(location_by_day_long, aes(x = day, y = votes, color = early_vote_center)) +
  geom_line() +
  labs(
    title = "Early Voting Trends by Location",
    x = "Day",
    y = "Number of Votes"
  ) +
  theme_minimal()
)



```


```{r}
### BY PRECINCT

precincts_by_day <- ev_raw|>
  group_by(precinct, county_name) |>
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

precincts_by_day <- precincts_by_day |>
  rowwise() |>
  mutate(
    first_day = `day1`,             
    last_day = `day8`,           
    max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[1],  
    second_max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[2],  
    follows_pattern = first_day %in% c(max_votes, second_max_votes) &
                      last_day %in% c(max_votes, second_max_votes) &
                      first_day != last_day)

summing_precincts <- precincts_by_day |>
  mutate(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))

precincts_percentage_follows_pattern <- precincts_by_day |>
  ungroup() |> 
  summarize(
    precincts_following = sum(follows_pattern),
    total_precincts = n(),
    percentage = (precincts_following / total_precincts) * 100
  )
```

```{r}
### BY COUNTY

counties_by_day_simple <- ev_raw|>
  group_by(county_name) |>
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

library(ggplot2)

counties_long <- counties_by_day_simple |>
  pivot_longer(cols = starts_with("day"), 
               names_to = "day", 
               values_to = "value")

counties_long <- counties_long |>
  group_by(county_name) |>
  mutate(last_day = if_else(day == last(day), TRUE, FALSE))

ggplot(counties_long, aes(x = day, y = value, color = county_name, group = county_name)) +
  geom_line() +
  geom_text(
    data = counties_long |> filter(last_day), 
    aes(label = county_name),
    hjust = -0.1, 
    vjust = 0.5, 
    size = 3
  ) +
  scale_x_discrete(labels = function(x) paste0("Day ", gsub("day", "", x))) + 
  labs(
    title = "Prince George's County and Montgomery County have most votes first and last days of early voting",
    x = "Day",
    y = "Number of Votes",
    color = "County"
  ) +
  theme_minimal()



counties_by_day <- counties_by_day_simple |>
 rowwise() |>
  mutate(
    first_day = `day1`,             
    last_day = `day8`,           
    max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[1],  
    second_max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[2],  
    follows_pattern = first_day %in% c(max_votes, second_max_votes) &
                     last_day %in% c(max_votes, second_max_votes) &
                      first_day != last_day)


summing_counties <- counties_by_day |>
  mutate(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))

counties_percentage_follows_pattern <- counties_by_day |>
  ungroup() |> 
  summarize(
    counties_following = sum(follows_pattern),
    total_counties = n(),
    percentage = (counties_following / total_counties) * 100
  )

### PERCENTAGE OF LOCATIONS IN COUNTY FOLLOWING PATTERN

locations_in_county_following <- location_by_day |>
  group_by(county_name) |>
  summarize(
    locations_following = sum(follows_pattern),
    total_locations = n(),
    percentage_following = (locations_following / total_locations) * 100
  )
```
```

### MOST NEWSWORTHY FINDING

We think our most newsworthy finding so far was found in Question 4 while looking at which counties had the closest margins between registered Republicans and registered Democrats who voted early. This question can help us identify split counties. We thought Kent County was particularly interesting because of how the same number of Democrats and Republicans voted early and because of how many unaffiliated voters voted early. The county seems small in comparison to other counties so we think looking into what factors impact how people in Kent County vote would be interesting. We also think comparing the margins of all of the other counties could make for an interesting story on its own. This would require making some visualizations and doing some more coding. 
