title: "Data Analysis Project"
author: "Marijke Friedman, Brendan Weissel, Sade Ajishegiri"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggthemes)
library(tidycensus)
```


```{r}
ev_raw <- read_csv("data/EV_RAW_DATA.csv") 

ev_raw <- ev_raw |>
  clean_names()
glimpse(ev_raw)
```

```{r}
#votes by day 
votes_by_day <- ev_raw |> 
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

#votes by party
votes_by_party <- ev_raw |>
  group_by(party_code) |> 
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE)) |> 
  arrange(party_code)

### votes by precinct 
votes_by_precinct <- ev_raw|>
  group_by(precinct)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(precinct)

### votes by gender
votes_by_gender <- ev_raw|>
  group_by(gender_code)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(gender_code)

### votes by age
votes_by_age <- ev_raw|>
  group_by(age_group)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(age_group)
```

### Our original questions:

1. What age group had the highest turn out for early voting?
2. What county had the highest turnout for early voting?
3. How does the 2024 early voting data compare to the 2020 early voting data?
4. What counties had the closest margin between registered Republicans and Democrats who voted early? Can we identify any swing counties from this?
5. How do early voting patterns vary among different demographic groups? 

###: Our new questions:

1. What precincts did younger age groups dominate voting early in? What do these precincts have in common?
2. Are there any precincts where certain demographic groups outperformed their Census data? For example, are there precincts were 65% of the early voters were women despite women only being 50% of the voting age population?
3. Which precincts had the most early voters identifying outside of male/female? How did they vote?
4. Which precincts have historically turned out more early voters than others? How did it compare this election?
5. Do most precincts follow the general pattern of having the most people vote on the first and last days of early voting? 

### How we've refined our questions:

Based on your feedback, we're refining our questions to focus more on precinct-level data instead of overall state data. We are looking more closely at patterns that stand out in certain precincts - taking note of the unusual and trying to find the reasons behind it.

### Describing any obvious limitations/flaws in our data - Why might we be unable to answer our previous questions? Describe the limits of the data in terms of the time frame, what is and is not included and any codes or other information we'll need to find definitions for, and what we've done to better define it. 

The data we have is just early voting data. It's interesting to analyze because early voters are likely to be more civic duty-minded. But the early voter data doesn't tell us that much about the overall electorate, so it could be helpful to bring in actual results data and more information about the overall electorate.

We'll likely use Census data so we can make more meaningful comparisons, plus voter registration data so we can look at voter turnout numbers. 

As far as defining some of the terms in our questions and within our data, in the votes_by_gender table, what does "NA" actually mean? Did voters in that category mark a box? Did they write in something other than male or female? Was there an option that said "prefer not to say"?
### Question 1: What voting location and or county did younger age groups dominate voting early in? What do these places have in common?

The state youth turnout % for early voting was 6.55%. 8 Counties had a higher %. 1. Saint Mary's (7.6%), 2. Howard (7.5%), 3. Charles (7.1%), 4. Calvert (7.1%), 5. Montgomery(6.9%), 6. Frederick(6.9%) 7. Prince George's (6.8%), 8. Washington (6.8%). The top counties tend to have a lower median age compared to other counties.

The top polling location for early youth vote is likely due to the proximity to the University of Maryland. 

The young vote percentage in PG County was 6.8% so the College Park Community Center out performed the county average by about 3 points with 10% being youth vote of the locations voting total.

```{r}
#total votes by county
total_votes_by_county <- ev_raw |>
  group_by(county_name) |>
  summarise(
    total_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
  )

#total votes by ages 18-24 by county
young_votes_by_county <- ev_raw |>
  filter(age_group == "18-24") |>
  group_by(county_name) |>
  summarise(
    young_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
  )

#calculating the voter share or percentage that young voters have in each county
young_voters_by_county_by_percent <- young_votes_by_county |>
  left_join(total_votes_by_county, by = "county_name") |>
  mutate(young_vote_share = young_votes / total_votes) |>
  arrange(desc(young_vote_share))

#total votes by early voting location
total_votes_by_location <- ev_raw |>
  group_by(early_vote_center, county_name) |>
  summarise(
    total_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  )

#total votes by ages 18-24 by early voting location
young_votes_by_location <- ev_raw |>
  filter(age_group == "18-24") |>
  group_by(early_vote_center) |>
  summarise(
    young_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  )

young_voters_by_location_by_percent <- young_votes_by_location |>
  left_join(total_votes_by_location, by = "early_vote_center") |>
  mutate(young_vote_share = young_votes / total_votes) |>
  arrange(desc(young_vote_share))

#Maryland total early youth vote %
total_young_votes <- ev_raw |>
  filter(age_group == "18-24") |>
  summarise(total_young_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE))

total_votes <- ev_raw |>
  summarise(total_votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE))

state_youth_turnout <- total_young_votes$total_young_votes / total_votes$total_votes * 100

state_youth_turnout


#18-24 on the census data
age_18_24 <- get_acs(
  geography = "county",
  variables = c(
    "B01001_007", # M 18-19
    "B01001_008", # M 20
    "B01001_009", # M 21
    "B01001_010", # M 22-24
    "B01001_031", # F 18-19
    "B01001_032", # F 20
    "B01001_033", # F 21
    "B01001_034"  # F 22-24
  ),
  state = "MD",
  year = 2022
) |>
  group_by(GEOID) |>
  summarise(youth_pop = sum(estimate))

#total population
total_pop <- get_acs(
  geography = "county",
  variables = "B01001_001", 
  state = "MD",
  year = 2022
)

#calculate percentage
youth_percent <- total_pop |>
  left_join(age_18_24, by = "GEOID") |>
  mutate(
    NAME = gsub(", Maryland", "", NAME),
    NAME = gsub(" County", "", NAME),
    NAME = case_when(
      NAME == "St. Mary's" ~ "Saint Mary's",
      NAME == "Baltimore" ~ "Baltimore County",
      NAME == "Baltimore city" ~ "Baltimore City",
      TRUE ~ NAME
    ),
    youth_percent = (youth_pop / estimate) * 100
  ) |>
  select(
    county_name = NAME,
    total_pop = estimate,
    youth_pop,
    youth_percent
  ) |>
  arrange(desc(youth_percent))

#combine census data with early voting data
Youth_voting_census_comparison <- young_voters_by_county_by_percent |>
  mutate(
    county_name = case_when(
      county_name == "St. Mary's" ~ "Saint Mary's",
      county_name == "Baltimore" ~ "Baltimore County",
      county_name == "Baltimore city" ~ "Baltimore City",
      TRUE ~ county_name
    )
  ) |>
  left_join(youth_percent, by = "county_name") |>
  mutate(
    young_vote_share = young_vote_share * 100,
  ) |>
  select(
    county_name,
    young_vote_share,
    census_youth_percent = youth_percent,
    total_votes,
    young_votes,
    census_youth_pop = youth_pop,
    census_total_pop = total_pop
  ) |>
  arrange(desc(census_youth_percent))

```

### Answering Question 2: Are there any precincts where certain demographic groups outperformed their Census data? For example, are there precincts were 65% of the early voters were women despite women only being 50% of the voting age population?

Chose to focus on gender for this one! - Did not look at N/A becasue we do that in Q3 and the number of them is so low

```{r}
# gender breakdown of early voters by county
ev_gender_breakdown <- ev_raw |>
 mutate(
   county_name = case_when(
     county_name == "St. Mary's" ~ "Saint Mary's",
     county_name == "Baltimore" ~ "Baltimore County",
     county_name == "Baltimore city" ~ "Baltimore City",
     TRUE ~ county_name
   )
 ) |>
   filter(gender_code != "N/A") |>  
 group_by(county_name, gender_code) |>
 summarise(
   votes_by_gender = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 group_by(county_name) |>
 mutate(
   total_votes = sum(votes_by_gender),
   voting_gender_percent = round((votes_by_gender/total_votes) * 100, 3)
 ) |>
 arrange(desc(voting_gender_percent))

# Census gender population data 
gender_population_data <- get_acs(
 geography = "county",
 variables = c(
   male = "B01001_002",    # Male total
   female = "B01001_026"   # Female total
 ),
 state = "MD",
 year = 2022
)

# Calculate gender percentages using census data
census_gender_percents <- gender_population_data |>
 mutate(
   NAME = gsub(", Maryland", "", NAME),
   NAME = gsub(" County", "", NAME),
   NAME = case_when(
     NAME == "St. Mary's" ~ "Saint Mary's",
     NAME == "Baltimore" ~ "Baltimore County",
     NAME == "Baltimore city" ~ "Baltimore City",
     TRUE ~ NAME
   )
 ) |>
 select(
   county_name = NAME,
   gender_code = variable,
   population = estimate
 ) |>
 group_by(county_name) |>
 mutate(
   total_pop = sum(population),
   census_gender_percent = round((population/total_pop) * 100, 3),
   gender_code = case_when(
     gender_code == "male" ~ "M",
     gender_code == "female" ~ "F"
   )
 ) |>
 arrange(county_name, gender_code)

# Compare census gender breakdown by county to early voting gender breakdown by county
gender_comparison <- ev_gender_breakdown |>
 left_join(census_gender_percents, by = c("county_name", "gender_code")) |>
 mutate(
   representation_difference = voting_gender_percent - census_gender_percent
 ) |>
 select(
   county_name,
   gender_code,
   voting_gender_percent,
   census_gender_percent,
   representation_difference,
   votes_by_gender,
   total_votes,
   population,
   total_pop
 ) |>
 arrange(desc(representation_difference))

```




### Answering Question 3:

```{r} 
ev_raw |> 
  group_by(gender_code) |> 
  count()
```
#Find percentage of nonbinary early voters:
```{r} 
(2777/(80405+77290+2777))*100
```
A: 1.7% of early voters were effectively non binary.

```{r}
# Find precincts with most N/A gender voters
nonbinary_precinct_analysis <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(precinct, county_name, early_vote_center) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 # Get total voters at each precinct for percentage
 left_join(
   ev_raw |>
     group_by(precinct) |>
     summarise(
       total_precinct_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
     ),
   by = "precinct"
 ) |>
 mutate(
   na_percent = round((na_voters/total_precinct_voters) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# Party breakdown for non-binary early voters
na_party_breakdown <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(party_code) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_na = sum(na_voters),
   party_percent = round((na_voters/total_na) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# County breakdown for non-binary early voters
na_county_breakdown <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(county_name) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_na = sum(na_voters),
   county_percent = round((na_voters/total_na) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# Using population data by county to see what county had the highest percent of non binary voters in relation to total county population
county_pop <- get_acs(
  geography = "county",
  variables = "B01001_001", 
  state = "MD",
  year = 2022
) |>
  mutate(
    NAME = gsub(", Maryland", "", NAME),
    NAME = gsub(" County", "", NAME),
    NAME = case_when(
      NAME == "St. Mary's" ~ "Saint Mary's",
      NAME == "Baltimore" ~ "Baltimore County",
      NAME == "Baltimore city" ~ "Baltimore City",
      TRUE ~ NAME
    )
  )

# Look at county breakdown of N/A gender voters with census context
na_county_census_breakdown <- ev_raw |>
  filter(is.na(gender_code)) |>
  group_by(county_name) |>
  summarise(
    na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(county_pop |> select(county_name = NAME, total_pop = estimate), 
            by = "county_name") |>
  mutate(
    total_na = sum(na_voters),
    county_percent = round((na_voters/total_na) * 100, 3),
    na_per_10k = round((na_voters/total_pop) * 10000, 3) 
  ) |>
  arrange(desc(na_per_10k))


```


### Answering Question 3:

```{r} 
ev_raw |> 
  group_by(gender_code) |> 
  count()
```
#Find percentage of nonbinary early voters:
```{r} 
(2777/(80405+77290+2777))*100
```
A: 1.7% of early voters were effectively non binary.

```{r}
# Find precincts with most non binary voters
nonbinary_precinct_analysis <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(precinct, county_name, early_vote_center) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 # Get total voters at each precinct for percentage
 left_join(
   ev_raw |>
     group_by(precinct) |>
     summarise(
       total_precinct_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
       .groups = "drop"
     ),
   by = "precinct"
 ) |>
 mutate(
   na_percent = round((na_voters/total_precinct_voters) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# Party breakdown for non-binary early voters
na_party_breakdown <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(party_code) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_na = sum(na_voters),
   party_percent = round((na_voters/total_na) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# County breakdown for non-binary early voters
na_county_breakdown <- ev_raw |>
 filter(is.na(gender_code)) |>
 group_by(county_name) |>
 summarise(
   na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 mutate(
   total_na = sum(na_voters),
   county_percent = round((na_voters/total_na) * 100, 3)
 ) |>
 arrange(desc(na_voters))

# Using population data by county to see what county had the highest percent of non binary voters in relation to total county population
county_pop <- get_acs(
  geography = "county",
  variables = "B01001_001", 
  state = "MD",
  year = 2022
) |>
  mutate(
    NAME = gsub(", Maryland", "", NAME),
    NAME = gsub(" County", "", NAME),
    NAME = case_when(
      NAME == "St. Mary's" ~ "Saint Mary's",
      NAME == "Baltimore" ~ "Baltimore County",
      NAME == "Baltimore city" ~ "Baltimore City",
      TRUE ~ NAME
    )
  )

# Look at county breakdown of N/A gender voters with census context
na_county_census_breakdown <- ev_raw |>
  filter(is.na(gender_code)) |>
  group_by(county_name) |>
  summarise(
    na_voters = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(county_pop |> select(county_name = NAME, total_pop = estimate), 
            by = "county_name") |>
  mutate(
    total_na = sum(na_voters),
    county_percent = round((na_voters/total_na) * 100, 3),
    na_per_10k = round((na_voters/total_pop) * 10000, 3) 
  ) |>
  arrange(desc(na_per_10k))


```

### Answering Question 4: What counties had the closest margin between registered Republicans and Democrats who voted early? Can we identify any swing counties from this?

Wicomico, Kent, Caroline, Talbot, Dorchester and Somerset counties had the closest margin between registered Republicans and Democrats who voted early. Kent County had exactly the same number of registered Republicans and Democrats who voted early. The number of unaffiliated voters in Kent County was not that much lower than the number of registered Republicans and Democrats who voted early. The total number of votes in Kent County during early voting was also quite low. I think the counties I listed are good starting places of where to look at swing counties, but all of the counties who had more Republicans than Democrats vote early are also places to look. Several of the counties with more Democrats than Republicans vote had much larger margins so I think it would be very safe to eliminate these from being swing counties. - NEEDS EDITS!!!

```{r}
party_by_county <- ev_raw |>
  group_by(county_name, party_code) |>
  summarise(
    votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
    .groups = "drop"
  ) |>
  group_by(county_name) |>
  mutate(
    total_votes = sum(votes),
    party_share = round((votes / total_votes) * 100, 1)
  ) |>
  summarise(
    total_votes = first(total_votes),
    votes_REP = sum(votes[party_code == "REP"], na.rm = TRUE),
    votes_DEM = sum(votes[party_code == "DEM"], na.rm = TRUE),
    votes_UNA = sum(votes[party_code == "UNA"], na.rm = TRUE),
    votes_OTH = sum(votes[party_code == "OTH"], na.rm = TRUE),
    votes_LIB = sum(votes[party_code == "LIB"], na.rm = TRUE),
    votes_GRN = sum(votes[party_code == "GRN"], na.rm = TRUE),
    votes_NLM = sum(votes[party_code == "NLM"], na.rm = TRUE),
    party_share_REP = sum(party_share[party_code == "REP"], na.rm = TRUE),
    party_share_DEM = sum(party_share[party_code == "DEM"], na.rm = TRUE),
    party_share_UNA = sum(party_share[party_code == "UNA"], na.rm = TRUE),
    party_share_OTH = sum(party_share[party_code == "OTH"], na.rm = TRUE),
    party_share_LIB = sum(party_share[party_code == "LIB"], na.rm = TRUE),
    party_share_GRN = sum(party_share[party_code == "GRN"], na.rm = TRUE),
    party_share_NLM = sum(party_share[party_code == "NLM"], na.rm = TRUE)
  ) |>
  mutate(
    margin_DEM_REP = abs(party_share_DEM - party_share_REP)
  ) |>
  arrange(margin_DEM_REP)


#now by precinct 
party_by_precinct <- ev_raw |>
 group_by(precinct, county_name, party_code) |>
 summarise(
   votes = sum(across(starts_with("day"), ~ .x), na.rm = TRUE),
   .groups = "drop"
 ) |>
 group_by(precinct, county_name) |>
 mutate(
   total_votes = sum(votes),
   party_share = round((votes / total_votes) * 100, 1)
 ) |>
 summarise(
   total_votes = first(total_votes),
   votes_REP = sum(votes[party_code == "REP"], na.rm = TRUE),
   votes_DEM = sum(votes[party_code == "DEM"], na.rm = TRUE),
   votes_UNA = sum(votes[party_code == "UNA"], na.rm = TRUE),
   votes_OTH = sum(votes[party_code == "OTH"], na.rm = TRUE),
   votes_LIB = sum(votes[party_code == "LIB"], na.rm = TRUE),
   votes_GRN = sum(votes[party_code == "GRN"], na.rm = TRUE),
   votes_NLM = sum(votes[party_code == "NLM"], na.rm = TRUE),
   party_share_REP = sum(party_share[party_code == "REP"], na.rm = TRUE),
   party_share_DEM = sum(party_share[party_code == "DEM"], na.rm = TRUE),
   party_share_UNA = sum(party_share[party_code == "UNA"], na.rm = TRUE),
   party_share_OTH = sum(party_share[party_code == "OTH"], na.rm = TRUE),
   party_share_LIB = sum(party_share[party_code == "LIB"], na.rm = TRUE),
   party_share_GRN = sum(party_share[party_code == "GRN"], na.rm = TRUE),
   party_share_NLM = sum(party_share[party_code == "NLM"], na.rm = TRUE),
   county_name = first(county_name)
 ) |>
 mutate(
   margin_DEM_REP = abs(party_share_DEM - party_share_REP),
   swing_precinct = case_when(
     margin_DEM_REP <= 5 ~ "Very Competitive (within 5%)",
     margin_DEM_REP <= 10 ~ "Competitive (5-10%)",
     TRUE ~ "Not Competitive (>10%)"
   )
 ) |>
 arrange(margin_DEM_REP)

swing_precinct_summary <- party_by_precinct |>
 group_by(swing_precinct) |>
 summarise(
   number_of_precincts = n(),
   .groups = "drop"
 ) |>
 mutate(
   percent_of_precincts = round((number_of_precincts/sum(number_of_precincts)) * 100, 1)
 ) |>
 arrange(desc(number_of_precincts))tes = REP + DEM + UNA + OTH + LIB + GRN + NLM)

```


### Answering Question 5: Do most precincts follow the general pattern of having the most people vote on the first AND last days of early voting?  Do most early voting locations follow the general pattern of having the most people vote on the first AND last days of early voting? Do most counties follow the general pattern of having the most people vote on the first AND last days of early voting?

No, most precincts do not follow the pattern of having the most people vote on the first AND last days of early voting. Only about 27% of precincts follow this pattern. For the second question, no most counties do not follow this pattern. The percentage is slightly higher at about 33%. One thing that I did notice through looking at this is that Prince George's County had significantly higher voters on the last day of early voting. Statewide, the first and last days of early voting had the most votes cast. If I look at this by polling locations, I can see that 36% of the ealry voting locations follow the pattern. I also can see that 14 of 24 counties have at least one polling location that follows the pattern. Among them are some of the counties with the most polling locations (Anne Arundel, Prince George's, Montgomery, Baltimore County, Baltimore City). These are some of the more populous regions in the state and some of those locations have more voters at them than others. This could be why statewide the first and last days of early voting had the most voters in Maryland, but this pattern was not followed when broken down by precinct, county, and polling location. 

### Note: I am going to return to make data visualizations and graphics for this question soon! I wanted to focus on starting on another question for this milestone. 

### Ideas for visualizations: Some line charts would be good and then we can talk about seeing each county (or voting location) using a facet_wrap in ggplot. A scatterplot could help illustrate the first day/last day pattern (and where it doesn't hold).


```{r}

### BY POLLING LOCATION

location_by_day <- ev_raw|>
  group_by(early_vote_center, county_name) |>
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

location_by_day <- location_by_day |>
  rowwise() |>
  mutate(
    first_day = `day1`,             
    last_day = `day8`,           
    max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[1],  
    second_max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[2],  
    follows_pattern = first_day %in% c(max_votes, second_max_votes) &
                      last_day %in% c(max_votes, second_max_votes) &
                      first_day != last_day)

summing_locations <- location_by_day |>
  mutate(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))

location_percentage_follows_pattern <- location_by_day |>
  ungroup() |> 
  summarize(
    location_following = sum(follows_pattern),
    total_location = n(),
    percentage = (location_following / total_location) * 100
  )

### BY PRECINCT

precincts_by_day <- ev_raw|>
  group_by(precinct, county_name) |>
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

precincts_by_day <- precincts_by_day |>
  rowwise() |>
  mutate(
    first_day = `day1`,             
    last_day = `day8`,           
    max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[1],  
    second_max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[2],  
    follows_pattern = first_day %in% c(max_votes, second_max_votes) &
                      last_day %in% c(max_votes, second_max_votes) &
                      first_day != last_day)

summing_precincts <- precincts_by_day |>
  mutate(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))

precincts_percentage_follows_pattern <- precincts_by_day |>
  ungroup() |> 
  summarize(
    precincts_following = sum(follows_pattern),
    total_precincts = n(),
    percentage = (precincts_following / total_precincts) * 100
  )

### BY COUNTY

counties_by_day <- ev_raw|>
  group_by(county_name) |>
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

counties_by_day <- counties_by_day |>
  rowwise() |>
  mutate(
    first_day = `day1`,             
    last_day = `day8`,           
    max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[1],  
    second_max_votes = sort(c_across(starts_with("day")), decreasing = TRUE)[2],  
    follows_pattern = first_day %in% c(max_votes, second_max_votes) &
                      last_day %in% c(max_votes, second_max_votes) &
                      first_day != last_day)


summing_counties <- counties_by_day |>
  mutate(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))

counties_percentage_follows_pattern <- counties_by_day |>
  ungroup() |> 
  summarize(
    counties_following = sum(follows_pattern),
    total_counties = n(),
    percentage = (counties_following / total_counties) * 100
  )

### PERCENTAGE OF LOCATIONS IN COUNTY FOLLOWING PATTERN

locations_in_county_following <- location_by_day |>
  group_by(county_name) |>
  summarize(
    locations_following = sum(follows_pattern),
    total_locations = n(),
    percentage_following = (locations_following / total_locations) * 100
  )
```

### MOST NEWSWORTHY FINDING

We think our most newsworthy finding so far was found in Question 4 while looking at which counties had the closest margins between registered Republicans and registered Democrats who voted early. This question can help us identify split counties. We thought Kent County was particularly interesting because of how the same number of Democrats and Republicans voted early and because of how many unaffiliated voters voted early. The county seems small in comparison to other counties so we think looking into what factors impact how people in Kent County vote would be interesting. We also think comparing the margins of all of the other counties could make for an interesting story on its own. This would require making some visualizations and doing some more coding. 
