---
title: "Class Reference"
author: "Marijke Friedman"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/Documents/GitHub/data_journalism_2024_fall")
```

### How to install and load the Tidyverse

This command loads and installs the Tidyverse libraries.

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to install and load R Markdown

This command allows me to use R Markdown files and is related to the Tidyverse.

```{r}
install.packages('rmarkdown')
```

### How to use assignment operators and do basic math in R

This command assigns and stores a value in a variable. I can use the variables to do math, including the multiplication I did below

```{r}
numberone <- 4 
numbertwo <- 2
numberone * numbertwo
```

### How to add the readr and dplyr libraries

This code loads the readr dplyr libraries in order to be able to read data. Loading library(tidyverse) is a shortcut to doing this. 

```{r}
library(readr)
library(dplyr)
```

### How to import data

This would help me import data if I had the file. The first part of the code is the name of a variable. The part following the assignment operator reads the data and stuffs it into the variable. read_rds only works when the Tidyverse has been loaded. It is a type of function

```{r}
maryland_winred_contributions <- read_rds("maryland_winred.rds")
```

### How to load a file from the internet

```{r}
maryland_winred_contributions <- read_rds("https://thescoop.org/files/maryland_winred.rds")

```

### How to easily view our data

The first function gives us a list of columns and their first few values and the data type in each column. The second function prints out the columns and the first six rows of data.

```{r}
glimpse(maryland_winred_contributions)
head(maryland_winred_contributions)

```

### How to do a specific action to a data set

This is taking your data set and doing a specific action to it. (data |> function)
 
A common pattern: data |> group_by(COLUMN NAME) |> summarize(VARIABLE NAME = AGGREGATE FUNCTION(COLUMN NAME))

The example below counts the number of contributions to each recipient.


```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n()
  )
```

### How to find the total, mean, median, minimum, and maximum
 
These functions can do math with our data.

```{r}
sum()
mean()
median()
min_amount()
max_amount()
```

### This function can arange our data from highest to lowest.

```{r}
arrange(desc())
```

### This function groups the data by city and finds the total number of contributions in each city. The arrange part of the function arranges the data in descending order based on which city had the most contributions. Taking out "desc" will arrange in ascending order. 

```{r}
wv_winred_contribs |>
  group_by(city) |>
  summarise(
    total_contribs = n()
  ) |>
  arrange(desc(total_contribs))
```

### This function finds what recipient got the largest total amount of money and how many contributions they received. It also shows me the average donation for each recipient. 

```{r}
wv_winred_contribs |>
  group_by(memo_text) |>
  summarise(
    total_contribs = n(),
    total_amount = sum(amount),
    average_amount = mean(amount
    )
  ) |>
  arrange(desc(total_amount))
```

#### This function produces  a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount. The first part of the function stores it as a variable object called `wv_summary` so that I can examine it with the data viewer. 

```{r}
wv_summary <- wv_winred_contribs |>
  group_by(memo_text) |>
   summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(total_amount))
```

### This function installs the lubridate package.

```{r}
library(lubridate)
```

### This function tells me how many new voters registered since July 21. The first part creates a new dataframe for those new registrants and filters it so it is only people since that date. Then it shows me what county had the most registrants since that date. 

```{r}
new_registrants <- md_voters_2024 |> filter(STATE_REGISTRATION_DATE >= "2024-07-22")

new_registrants |>
  group_by(COUNTY) |>
  summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))

glimpse(new_registrants)
```

### This function mutates the original dataframe to add a new column called "week." This is all through lubridate. The second part of the function groups and summarizes the information by week and then arranges the weeks in order by number of new registrants per week. 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(week = floor_date(STATE_REGISTRATION_DATE, "week")) 
md_voters_2024 |>
   group_by(week) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))
```

### This function does the same thing as the last one and just does it by month instead of week using lubridate. 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(month = floor_date(STATE_REGISTRATION_DATE, "month")) 
md_voters_2024 |>
   group_by(month) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))
```

### The below code does a lot. First, it mutates the original dataframe to add a column called "region." It also uses case_when to assign a region to each county. The second part arranges the regions in order based on the number of new registrants. The last part groups and summarizes the new registrants by political party and arranges them in order of which party had the most new registrants. 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(
    region = case_when(
        COUNTY == "Garrett" ~ "Western Maryland",
        COUNTY == "Allegany" ~ "Western Maryland",
        COUNTY == "Washington" ~ "Western Maryland",
        COUNTY == "Frederick" ~ "Capital Region",
        COUNTY == "Montgomery" ~ "Capital Region",
        COUNTY == "Prince George's" ~ "Capital Region",
        COUNTY == "Anne Arundel" ~ "Central Maryland",
        COUNTY == "Baltimore City" ~ "Central Maryland",
        COUNTY == "Baltimore County" ~ "Central Maryland",
        COUNTY == "Carroll" ~ "Central Maryland",
        COUNTY == "Harford" ~ "Central Maryland",
        COUNTY == "Howard" ~ "Central Maryland",
        COUNTY == "Calvert" ~ "Southern Maryland",
        COUNTY == "Charles" ~ "Southern Maryland",
        COUNTY == "Saint Mary's" ~ "Southern Maryland",
        COUNTY == "Kent" ~ "Eastern Shore",
        COUNTY == "Queen Anne's" ~ "Eastern Shore",
        COUNTY == "Talbot" ~ "Eastern Shore",
        COUNTY == "Caroline" ~ "Eastern Shore",
        COUNTY == "Dorchester" ~ "Eastern Shore",
        COUNTY == "Somerset" ~ "Eastern Shore",
        COUNTY == "Worcester" ~ "Eastern Shore",
        COUNTY == "Cecil" ~ "Eastern Shore",
        COUNTY == "Kent" ~ "Eastern Shore"
        )
  )
md_voters_2024 |>
   group_by(region) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))

md_voters_2024 |>
   group_by(PARTY) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))
```

### This reads and loads only the first ten rows of data.

```{r}
texas_precinct_20 <- read_csv("data/tx_precinct_2020.csv", guess_max=10)
```

### This checks for problems. It gives us how many rows have problems and sort of shows us what the problems are. 

```{r}
problems(texas_precinct_20)
```

### This allows us to find where the data type is a problem. It read in as a number (chr), even though there clearly are numbers in it judging from our initial inspection. Here's why: the original file has a single value in that column that is "5+".

```{r}
texas_precinct_20 |> filter(election_day == "5+")

```

### This fixes the above value and changes the "election_day" field data type. 

### ".default = election_day leaves all the others unchanged 

### as.numeric converts column type into number

```{r}
texas_precinct_20 <- texas_precinct_20 |>
  mutate(election_day = case_when(
    election_day == '5+' ~ '49',
    .default = election_day
  ))

texas_precinct_20 <- texas_precinct_20 |> mutate(election_day = as.numeric(election_day))

glimpse(texas_precinct_20)
```

### This creates a new column called "registration_month" that extracts the month and year from the "registr_dt" column. It groups by the new "registration_month" column, count, and sort by "registration_month". 

```{r}
yadkin_voters |> 
  mutate(registration_month = floor_date(registr_dt, "month")) |>
  group_by(registration_month) |> 
   summarise(
    count=n()
  ) |>
  arrange(registration_month)
```

### This installs and loads the janitor package.

```{r}
install.packages('janitor')
library(janitor)
```

### This cleans the dataset and assigns the cleaned dataset a new name and then displays it.

```{r}
# cleaning function
cleaned_conowingo <- conowingo |>
  clean_names()

# display the cleaned dataset
cleaned_conowingo
```

### This cleans and renames column names.

### Rename and clean name only work on column names. The rename comes first in the function.

```{r}
cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber)

# display the cleaned dataset
cleaned_conowingo
```

### This changes the data type to dbl instead of chr so we can do math with it. 

```{r}
# cleaning function
cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount))
  

# display the cleaned dataset
cleaned_conowingo
```

### This checks for duplicate rows.

```{r}
cleaned_conowingo |> 
  get_dupes()
```

### This gets rid of duplicate rows.

```{r}
# cleaning function
cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount)) |> 
  distinct()
```

### This standardizes zipcodes to 5 numbers.

```{r}
# cleaning function
cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount)) |> 
  distinct() |>
  mutate(zip = str_sub(zip, start=1L, end=5L))
  

# display the cleaned dataset
cleaned_conowingo

```

### This standardizes capitalization of the cities.
```{r}
# cleaning function
cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount)) |> 
  distinct() |>
  mutate(zip = str_sub(zip, start=1L, end=5L)) |>
  mutate(city = str_to_title(city))
  

# display the cleaned dataset
cleaned_conowingo
```

### This fixes some misspellings.

```{r}
# cleaning function
cleaned_conowingo <- conowingo |>
  clean_names() |> 
  rename(linenumber = x1_linenumber) |> 
  mutate(amount = as.numeric(amount)) |> 
  distinct() |>
  mutate(zip = str_sub(zip, start=1L, end=5L)) |>
  mutate(city = str_to_title(city)) |>
  mutate(city = case_when(
    city == "Conowing" ~ "Conowingo",
    TRUE ~ city
  ))

# display the cleaned dataset
cleaned_conowingo

```

### This creates a new dataframe that is just by contributor cities and arranges them. Then it cleans typos in that datagrame and makes sure the rows are distinct. Then it organizes again. 

```{r}
contribs_by_city <- md_contribs |>
  group_by(contributor_city) |>
  summarise(
    total_contribs = n()
  ) |>
  arrange((contributor_city))

cleaned_typos <- md_contribs |>
  clean_names() |> 
  mutate(contributor_city = case_when(
    contributor_city == "BALRIMORE" ~ "BALTIMORE",
    contributor_city == "BLATIMORE" ~ "BALTIMORE",
    contributor_city == "ANAPOLIS" ~ "ANNAPOLIS",
    contributor_city == "POTAMIC" ~ "POTOMAC",
    contributor_city == "SILVER SPRINGS" ~ "SILVER SPRING",
    contributor_city == "4615 NORTH PARK AVE APT. 1410" ~ "CHEVY CHASE",
    TRUE ~ contributor_city)) |> 
  distinct()

cleaned_contribs_by_city <- cleaned_typos |>
  group_by(contributor_city) |>
  summarise(
    total_contribs = n()
  ) |>
  arrange((contributor_city))
```

### This converts to a 5 digit zipcode. Then it sorts and groups by zipcode and contibutor and also arranges. Then it looks at a new dataframe filtered for the top zipcode and groups by committee name. 

```{r}
cleaned_typos <- cleaned_typos |>
  mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))
  
sort_by_zip <- cleaned_typos |>
    group_by(contributor_zip) |>
  summarise(
    total_amount = sum(contribution_receipt_amount
    )
  ) |>
  arrange(desc(total_amount))

zip_recipients <- cleaned_typos |> filter(contributor_zip == "20817")
 
zip_recipients <- zip_recipients|>
  group_by(committee_name) |>
  summarise(
    total_amount = sum(contribution_receipt_amount
    )
  ) |>
  arrange(desc(total_amount))
  
```

### This makes a new top donors dataframe and groups by contributor name and contributor city. It also sums the contribution amount and then arranges. 

```{r}
top_donors <-cleaned_typos |>
  group_by(contributor_name,contributor_city)|>
  summarise(total_amount = sum(contribution_receipt_amount))|>
  arrange(desc(total_amount))

```

### This uses data that has been cleaned in open refine to find the largest amounts of spending on categories by both campaigns.

```{r}
 new_senate_expenses |>
  group_by(purpose_clean) |>
  summarise(
    total_contribs = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))

hogan_expenses <- new_senate_expenses |> filter(candidate == "Hogan") |>
 group_by(purpose_clean) |>
  summarise(
    total_contribs = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))
 
alsobrooks_expenses <- new_senate_expenses |> filter(candidate == "Alsobrooks") |>
 group_by(purpose_clean) |>
  summarise(
    total_contribs = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))
```

### This allows us to pick one of the top categories for each campaign and shows how that spending has changed over time.

```{r}
top_category <- alsobrooks_expenses |>
  slice (1)|>
  pull(purpose_clean)

alsobrooks_by_date <- new_senate_expenses|>
  filter(candidate=="Alsobrooks", purpose_clean == top_category)|>
   mutate(week = floor_date(date, "week")) |>
    group_by(week) |>
    summarise(
    total_amount = sum(amount))|>
    arrange(week)

top_category <- hogan_expenses |>
  slice (1)|>
  pull(purpose_clean)

hogan_by_date <- new_senate_expenses|>
  filter(candidate=="Hogan", purpose_clean == top_category)|>
   mutate(week = floor_date(date, "week")) |>
    group_by(week) |>
    summarise(
    total_amount = sum(amount))|>
    arrange(week)
```

### This shows us which committee has the larger payroll or salary spending so far and how many unique recipients of payroll/staff spending there are for each candidate. 

```{r}
alsobrooks_salary <- new_senate_expenses |>
  filter(candidate == "Alsobrooks", purpose_clean == "Salary")

hogan_salary <- new_senate_expenses |>
  filter(candidate == "Hogan", purpose_clean == "Salary")

total_alsobrooks_salary <- alsobrooks_salary |>
  summarise(total_amount = sum(amount))

total_hogan_salary <- hogan_salary |>
  summarise(total_amount = sum(amount))

unique_alsobrooks_recipients <- alsobrooks_salary |>
  summarise(unique_recipients = n_distinct(first_name))

unique_hogan_recipients <- hogan_salary |>
  summarise(unique_recipients = n_distinct(organization_name))
```

### This shows us how to combine the  files into a single data frame using the bind_rows() function, along with list().

```{r}
# bind_rows with list
county_voters_combined <- bind_rows(list(county_voters_2016, county_voters_2018, county_voters_2020, county_voters_2022, county_voters_2024))
View(county_voters_combined)
```

### This puts the Maryland population data and voter registration data together through a join. There are different kinds of joins. It’s better if you think of two tables sitting next to each other. A left_join takes all the records from the left table and only the records that match in the right one. A right_join does the same thing. An inner_join takes only the records where they are equal. There’s one other join – a full_join which returns all rows of both, regardless of if there’s a match – but I’ve never once had a use for a full join.

### In the best-case scenario, the two tables we want to join share a common column. In this case, both of our tables have a column called county that has the same characteristics: values in both look identical, including how they distinguish Baltimore City from Baltimore County. This is important, because joins work on exact matches.

### We can do this join multiple ways and get a similar result. We can put the population file on the left and the registration data on the right and use a left join to get them all together. And we use join_by() to join by the correct columns. I’m going to count the rows at the end. The reason I’m doing this is important: Rule 1 in joining data is having an idea of what you are expecting to get. So with a left join with population on the left, I have 24 rows, so I expect to get 24 rows when I’m done.

```{r}
maryland_population |> left_join(county_voters_2020, join_by(COUNTY)) |> nrow()

```

### This assigns joined data to a new dataframe.

```{r}
maryland_population_with_voters <- maryland_population |> left_join(county_voters_2020)

```

### This uses a left_join to make our demographic data available. This time we’ll need to specify the two fields to join because they do not have identical names. We’ll use COUNTY from our population data and NAME from the demographic data, and the order matters - the first column is from the dataframe you name first.

### need to do: first data frame, first join column, second dataframe, second join column

```{r}
maryland_population_with_voters_and_demographics <- maryland_population_with_voters |> left_join(maryland_demographics, join_by(COUNTY==NAME))
```

### This uses mutate to find the county with the highest 18+ Black population as a percentage of all population 18 and over and also the percentage of Democrats in that county. 

```{r}
maryland_population_with_voters_and_demographics |>
  mutate(pct_black_18_plus = (pop_black/pop_18_over)*100, pct_dems = (DEM/TOTAL)*100) |> 
  arrange(desc(pct_black_18_plus)) |> 
  select(COUNTY, pct_black_18_plus, pct_dems)
```

### This changes the zipcode from a number to a character because zipcodes should never be numbers. 

```{r}
maryland_zcta <- maryland_zcta |> mutate(across(ZCTA5N, as.character))
glimpse(maryland_zcta)
```

### Look in Lab 5 for more helps with joins. 