---
title: "Class Reference"
author: "Marijke Friedman"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/Documents/GitHub/data_journalism_2024_fall")
```

### How to install and load the Tidyverse

This command loads and installs the Tidyverse libraries.

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to install and load R Markdown

This command allows me to use R Markdown files and is related to the Tidyverse.

```{r}
install.packages('rmarkdown')
```

### How to use assignment operators and do basic math in R

This command assigns and stores a value in a variable. I can use the variables to do math, including the multiplication I did below

```{r}
numberone <- 4 
numbertwo <- 2
numberone * numbertwo
```

### How to add the readr and dplyr libraries

This code loads the readr dplyr libraries in order to be able to read data. Loading library(tidyverse) is a shortcut to doing this. 

```{r}
library(readr)
library(dplyr)
```

### How to import data

This would help me import data if I had the file. The first part of the code is the name of a variable. The part following the assignment operator reads the data and stuffs it into the variable. read_rds only works when the Tidyverse has been loaded. It is a type of function

```{r}
maryland_winred_contributions <- read_rds("maryland_winred.rds")
```

### How to load a file from the internet

```{r}
maryland_winred_contributions <- read_rds("https://thescoop.org/files/maryland_winred.rds")

```

### How to easily view our data

The first function gives us a list of columns and their first few values and the data type in each column. The second function prints out the columns and the first six rows of data.

```{r}
glimpse(maryland_winred_contributions)
head(maryland_winred_contributions)

```

### How to do a specific action to a data set

This is taking your data set and doing a specific action to it. (data |> function)
 
A common pattern: data |> group_by(COLUMN NAME) |> summarize(VARIABLE NAME = AGGREGATE FUNCTION(COLUMN NAME))

The example below counts the number of contributions to each recipient.


```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n()
  )
```

### How to find the total, mean, median, minimum, and maximum
 
These functions can do math with our data.

```{r}
sum()
mean()
median()
min_amount()
max_amount()
```

### This function can arange our data from highest to lowest.

```{r}
arrange(desc())
```

### This function groups the data by city and finds the total number of contributions in each city. The arrange part of the function arranges the data in descending order based on which city had the most contributions. Taking out "desc" will arrange in ascending order. 

```{r}
wv_winred_contribs |>
  group_by(city) |>
  summarise(
    total_contribs = n()
  ) |>
  arrange(desc(total_contribs))
```

### This function finds what recipient got the largest total amount of money and how many contributions they received. It also shows me the average donation for each recipient. 

```{r}
wv_winred_contribs |>
  group_by(memo_text) |>
  summarise(
    total_contribs = n(),
    total_amount = sum(amount),
    average_amount = mean(amount
    )
  ) |>
  arrange(desc(total_amount))
```

#### This function produces  a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount. The first part of the function stores it as a variable object called `wv_summary` so that I can examine it with the data viewer. 

```{r}
wv_summary <- wv_winred_contribs |>
  group_by(memo_text) |>
   summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(total_amount))
```

### This function installs the lubridate package.

```{r}
library(lubridate)
```

### This function tells me how many new voters registered since July 21. The first part creates a new dataframe for those new registrants and filters it so it is only people since that date. Then it shows me what county had the most registrants since that date. 

```{r}
new_registrants <- md_voters_2024 |> filter(STATE_REGISTRATION_DATE >= "2024-07-22")

new_registrants |>
  group_by(COUNTY) |>
  summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))

glimpse(new_registrants)
```

### This function mutates the original dataframe to add a new column called "week." This is all through lubridate. The second part of the function groups and summarizes the information by week and then arranges the weeks in order by number of new registrants per week. 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(week = floor_date(STATE_REGISTRATION_DATE, "week")) 
md_voters_2024 |>
   group_by(week) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))
```

### This function does the same thing as the last one and just does it by month instead of week using lubridate. 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(month = floor_date(STATE_REGISTRATION_DATE, "month")) 
md_voters_2024 |>
   group_by(month) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))
```

### The below code does a lot. First, it mutates the original dataframe to add a column called "region." It also uses case_when to assign a region to each county. The second part arranges the regions in order based on the number of new registrants. The last part groups and summarizes the new registrants by political party and arranges them in order of which party had the most new registrants. 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(
    region = case_when(
        COUNTY == "Garrett" ~ "Western Maryland",
        COUNTY == "Allegany" ~ "Western Maryland",
        COUNTY == "Washington" ~ "Western Maryland",
        COUNTY == "Frederick" ~ "Capital Region",
        COUNTY == "Montgomery" ~ "Capital Region",
        COUNTY == "Prince George's" ~ "Capital Region",
        COUNTY == "Anne Arundel" ~ "Central Maryland",
        COUNTY == "Baltimore City" ~ "Central Maryland",
        COUNTY == "Baltimore County" ~ "Central Maryland",
        COUNTY == "Carroll" ~ "Central Maryland",
        COUNTY == "Harford" ~ "Central Maryland",
        COUNTY == "Howard" ~ "Central Maryland",
        COUNTY == "Calvert" ~ "Southern Maryland",
        COUNTY == "Charles" ~ "Southern Maryland",
        COUNTY == "Saint Mary's" ~ "Southern Maryland",
        COUNTY == "Kent" ~ "Eastern Shore",
        COUNTY == "Queen Anne's" ~ "Eastern Shore",
        COUNTY == "Talbot" ~ "Eastern Shore",
        COUNTY == "Caroline" ~ "Eastern Shore",
        COUNTY == "Dorchester" ~ "Eastern Shore",
        COUNTY == "Somerset" ~ "Eastern Shore",
        COUNTY == "Worcester" ~ "Eastern Shore",
        COUNTY == "Cecil" ~ "Eastern Shore",
        COUNTY == "Kent" ~ "Eastern Shore"
        )
  )
md_voters_2024 |>
   group_by(region) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))

md_voters_2024 |>
   group_by(PARTY) |>
   summarise(
    number_of_registrants = n()
  ) |>
  arrange(desc(number_of_registrants))
```

