---
title: "Class Reference"
author: "Marijke Friedman"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/Documents/GitHub/data_journalism_2024_fall")
```

### How to install and load the Tidyverse

This command loads and installs the Tidyverse libraries.

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to install and load R Markdown

This command allows me to use R Markdown files and is related to the Tidyverse.

```{r}
install.packages('rmarkdown')
```

### How to use assignment operators and do basic math in R

This command assigns and stores a value in a variable. I can use the variables to do math, including the multiplication I did below

```{r}
numberone <- 4 
numbertwo <- 2
numberone * numbertwo
```

### How to add the readr and dplyr libraries

This code loads the readr dplyr libraries in order to be able to read data. Loading library(tidyverse) is a shortcut to doing this. 

```{r}
library(readr)
library(dplyr)
```

### How to import data

This would help me import data if I had the file. The first part of the code is the name of a variable. The part following the assignment operator reads the data and stuffs it into the variable. read_rds only works when the Tidyverse has been loaded. It is a type of function

```{r}
maryland_winred_contributions <- read_rds("maryland_winred.rds")
```

### How to load a file from the internet

```{r}
maryland_winred_contributions <- read_rds("https://thescoop.org/files/maryland_winred.rds")

```

### How to easily view our data

The first function gives us a list of columns and their first few values and the data type in each column. The second function prints out the columns and the first six rows of data.

```{r}
glimpse(maryland_winred_contributions)
head(maryland_winred_contributions)

```

### How to do a specific action to a data set

This is taking your data set and doing a specific action to it. (data |> function)
 
A common pattern: data |> group_by(COLUMN NAME) |> summarize(VARIABLE NAME = AGGREGATE FUNCTION(COLUMN NAME))

The example below counts the number of contributions to each recipient.


```{r}
maryland_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n()
  )
```

### How to find the total, mean, median, minimum, and maximum
 
These functions can do math with our data.

```{r}
sum()
mean()
median()
min_amount()
max_amount()
```

### This function can arange our data from highest to lowest.

```{r}
arrange(desc())
```