---
title: "lab_01"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia, which is in a file called `lab_01.rds`

**Task** Using the example from the pre_lab, write code to load the RDS file and assign it to a good variable name.

```{r}
west_virginia_winred_contributions <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.

**Q1** How many rows and columns does this dataframe have?

**A1** This dataframe has 40,441 rows and 24 columns.

```{r}
glimpse(west_virginia_winred_contributions)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?

**A2** Charleston, West Virginia. There were 3226 contributions. 

```{r}
west_virginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n()
  ) |>
arrange(desc(count_contribs))
```

**Q3** Which West Virginia cities accounted for the fewest contributions? How many of them, roughly speaking?

**A3**Around 60 cities in West Virginia tied with only 1 contribution each.  

```{r}
west_virginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n()
  ) |>
arrange(count_contribs)
```

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.

**A4**Charleston is the capital of West Virginia so probably has a larger population than a lot of these other cities. Also, since the state government is primary located in this capital city, its population may be more politically aware and active. The other cities are smaller and not the capital city.

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?

**A5** The Trump National Committee JFC, Inc. It received 14063 donations and $463,944.40. The average donation was $32.99.

```{r}
west_virginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set. You do not need to write any code for this question, so no codeblock is provided.

**A6** That top recipient is Trump's Joint Fundraising Committee with the RNC. This is popular because Trump is a popular candidate in West Virginia. He won in West Virginia in 2020.

**Q7**. Using the answer from A5 as a starting point, produce a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**I noticed that the largest average for campaign contributions was for Guthrie for Congress at $3,300. Even though the Trump campaign got the most contributions and money total, its average contribution was a lot smaller. It could be interesting to look at why the Guthrie campaign and others draw in larger contributions on average. There are also big gaps between the maximum amounts of contribution. The Trump National Committee JFC had a $10,000 maximum contribution and and everything else was a lot less. The refund of earmarked contribution was also something that jumped out. I am curious why so many contributions were refunded and think that is something I would want to explore and possibly include in a news story. 

```{r}
'wv_summary' <- west_virginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange(desc(total_amount))
```
