---
title: "Unaffiliated Voters"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen=999)

library(tidyverse)
library(janitor)
```

```{r}
# 2024 registrations
registrations_2024 <- read_csv("october_data/oct_24.csv") |>
  rename(DEM24=DEM, REP24=REP, NLM24=NLM, GRN24=GRN, LIB24=LIB, UNAF24=UNAF, OTH24=OTH, TOTAL24=TOTAL)
  
```

```{r}
# 2022 registrations
registrations_2022 <- read_csv("october_data/oct_22.csv")|>
 rename(DEM22=DEM, REP22=REP, GRN22=GRN, LIB22=LIB, WCP22=WCP, UNAF22=UNAF, OTH22=OTH, TOTAL22=TOTAL)

```

```{r}
# 2020 registrations
registrations_2020 <- read_csv("october_data/oct_20.csv")|>
 rename(DEM20=DEM, REP20=REP, GRN20=GRN, LIB20=LIB, WCP20=WCP, UNAF20=UNAF, OTH20=OTH, TOTAL20=TOTAL, BAR20=BAR)
```

```{r}
# 2018 registrations
registrations_2018 <- read_csv("october_data/oct_18.csv")|>
  rename(DEM18=DEM, REP18=REP, GRN18=GRN, LIB18=LIB, UNAF18=UNAF, OTH18=OTH, TOTAL18=TOTAL)
```

```{r}
# 2016 registrations
registrations_2016 <- read_csv("october_data/oct_16.csv")|>
  rename(DEM16=DEM, REP16=REP, GRN16=GRN, LIB16=LIB, UNAF16=UNAF, OTH16=OTH, TOTAL16=TOTAL)
```

```{r}
# 2014 registrations
registrations_2014 <- read_csv("october_data/oct_14.csv")|>
  rename(DEM14=DEM, REP14=REP, GRN14=GRN, LIB14=LIB, UNAF14=UNAF, OTH14=OTH, TOTAL14=TOTAL)
```

```{r}
# 2012 registrations
registrations_2012 <- read_csv("october_data/oct_12.csv")|>
  rename(DEM12=DEM, REP12=REP, GRN12=GRN, LIB12=LIB, AmE12=AmE, UNAF12=UNAF, OTH12=OTH, TOTAL12=TOTAL)
```

```{r}
# 2010 registrations
registrations_2010 <- read_csv("october_data/oct_10.csv")|>
  rename(DEM10=DEM, REP10=REP, GRN10=GRN, LIB10=LIB, UNAF10=UNAF, OTH10=OTH, TOTAL10=TOTAL)
```

```{r}
# 2008 registrations
registrations_2008 <- read_csv("october_data/oct_08.csv")|>
  rename(DEM08=DEM, REP08=REP, GRN08=GRN, LIB08=LIB, UNAF08=UNAF, OTH08=OTH, TOTAL08=TOTAL, CON08=CON, IND08=IND)
```

```{r}
# 2006 registrations
registrations_2006 <- read_csv("october_data/fixed_oct_06.csv")|>
  rename(DEM06=DEM, REP06=REP, GRN06=GRN, CONST06=CONST, LIB06=LIB, POP06=POP, UNAF06=UNAF, OTH06=OTH, TOTAL06=TOTAL)
```

```{r}
# 2004 registrations
registrations_2004 <- read_csv("october_data/oct_04.csv")|>
 rename(DEM04=DEM, REP04=REP, GRN04=GRN, LIB04=LIB, UNAF04=UNAF, OTH04=OTH, TOTAL04=TOTAL, CONST04=CONST, POP04=POP)
```

### This shows was a -.94% change in Democratic voters, a 1.9% increase in Republican voters and a 10.2% increase in unaffiliated voters between September 2022 and September 2024. I also see that more than 20% of Maryland's voters are registered as unaffiliated. 

```{r}
registrations_2024 <- registrations_2024 |> 
  mutate(pct_DEM24 = (DEM24/TOTAL24)*100, pct_REP24 = (REP24/TOTAL24)*100, pct_UNAF24 = (UNAF24/TOTAL24)*100)

registrations_2022 <- registrations_2022 |> 
  mutate(pct_DEM22 = (DEM22/TOTAL22)*100, pct_REP22 = (REP22/TOTAL22)*100, pct_UNAF22 = (UNAF22/TOTAL22)*100)

compare_22_24 <- registrations_2024 |> 
  left_join(registrations_2022, by='COUNTY') |> 
  mutate(DEM_pct_change = (DEM24-DEM22)/DEM22*100, REP_pct_change = (REP24-REP22)/REP22*100, UNAF_pct_change = (UNAF24-UNAF22)/UNAF22*100) |> 
  arrange(desc(UNAF_pct_change))
```

### This shows there was an 31.56% increase in registered Democrats, a 13.6% increaqse in registered Republicans and a 117.8% increase in registered unaffiliated voters between 2004 and 2024.

```{r}
registrations_2024 <- registrations_2024 |> 
  mutate(pct_DEM24 = (DEM24/TOTAL24)*100, pct_REP24 = (REP24/TOTAL24)*100, pct_UNAF24 = (UNAF24/TOTAL24)*100)

registrations_2004 <- registrations_2004 |> 
  mutate(pct_DEM04 = (DEM04/TOTAL04)*100, pct_REP04 = (REP04/TOTAL04)*100, pct_UNAF04 = (UNAF04/TOTAL04)*100)

compare_04_24 <- registrations_2024 |> 
  left_join(registrations_2004, by='COUNTY') |> 
  mutate(DEM_pct_change = (DEM24-DEM04)/DEM04*100, REP_pct_change = (REP24-REP04)/REP04*100, UNAF_pct_change = (UNAF24-UNAF04)/UNAF04*100) |> 
  arrange(UNAF_pct_change)
```

### Getting percentages for the other dataframes!

```{r}
registrations_2016 <- registrations_2016 |> 
  mutate(pct_DEM16 = (DEM16/TOTAL16)*100, pct_REP16 = (REP16/TOTAL16)*100, pct_UNAF16 = (UNAF16/TOTAL16)*100)

registrations_2018 <- registrations_2018 |> 
  mutate(pct_DEM18 = (DEM18/TOTAL18)*100, pct_REP18 = (REP18/TOTAL18)*100, pct_UNAF18 = (UNAF18/TOTAL18)*100)

registrations_2020 <- registrations_2020 |> 
  mutate(pct_DEM20 = (DEM20/TOTAL20)*100, pct_REP20 = (REP20/TOTAL20)*100, pct_UNAF20 = (UNAF20/TOTAL20)*100)

registrations_2014 <- registrations_2014 |> 
  mutate(pct_DEM14 = (DEM14/TOTAL14)*100, pct_REP14 = (REP14/TOTAL14)*100, pct_UNAF14 = (UNAF14/TOTAL14)*100)

registrations_2012 <- registrations_2012 |> 
  mutate(pct_DEM12 = (DEM12/TOTAL12)*100, pct_REP12 = (REP12/TOTAL12)*100, pct_UNAF12 = (UNAF12/TOTAL12)*100)

registrations_2010 <- registrations_2010 |> 
  mutate(pct_DEM10 = (DEM10/TOTAL10)*100, pct_REP10 = (REP10/TOTAL10)*100, pct_UNAF10 = (UNAF10/TOTAL10)*100)

registrations_2008 <- registrations_2008 |> 
  mutate(pct_DEM08 = (DEM08/TOTAL08)*100, pct_REP08 = (REP08/TOTAL08)*100, pct_UNAF08 = (UNAF08/TOTAL08)*100)

registrations_2006 <- registrations_2006 |> 
  mutate(pct_DEM06 = (DEM06/TOTAL06)*100, pct_REP06 = (REP06/TOTAL06)*100, pct_UNAF06= (UNAF06/TOTAL06)*100)
```

### Trying to do one big merge for comparison!

### Unaffiliated percent change starts being the largest (in the limited years I am looking at) starting at 2016-2018 percent change.

```{r}
compare_all <- registrations_2024 |> 
  left_join(registrations_2004, by='COUNTY') |> 
  left_join(registrations_2006, by='COUNTY') |> 
  left_join(registrations_2008, by='COUNTY') |> 
  left_join(registrations_2010, by='COUNTY') |> 
  left_join(registrations_2012, by='COUNTY') |> 
  left_join(registrations_2014, by='COUNTY') |> 
  left_join(registrations_2016, by='COUNTY') |> 
  left_join(registrations_2018, by='COUNTY') |> 
  left_join(registrations_2020, by='COUNTY') |> 
  left_join(registrations_2022, by='COUNTY') |> 
  mutate(
    #2004-2006 pct_change
    DEM0406_pct_change = (DEM06-DEM04)/DEM06*100, REP0406_pct_change = (REP06-REP04)/REP06*100, UNAF0406_pct_change = (UNAF06-UNAF04)/UNAF04*100,
    #2006-2008 pct_change
    DEM0608_pct_change = (DEM08-DEM06)/DEM06*100, REP0608_pct_change = (REP08-REP06)/REP06*100, UNAF0608_pct_change = (UNAF08-UNAF06)/UNAF06*100,
     #2008-2010 pct_change
    DEM0810_pct_change = (DEM10-DEM08)/DEM08*100, REP0810_pct_change = (REP10-REP08)/REP08*100, UNAF0810_pct_change = (UNAF10-UNAF08)/UNAF08*100,
    #2010-2012 pct_change
    DEM1012_pct_change = (DEM12-DEM10)/DEM10*100, REP1012_pct_change = (REP12-REP10)/REP10*100, UNAF1012_pct_change = (UNAF12-UNAF10)/UNAF10*100,
    #2012-2014 pct_change
    DEM1214_pct_change = (DEM14-DEM12)/DEM12*100, REP1214_pct_change = (REP14-REP12)/REP12*100, UNAF1214_pct_change = (UNAF14-UNAF12)/UNAF12*100,
    #2014-2016 pct_change
    DEM1416_pct_change = (DEM16-DEM14)/DEM14*100, REP1416_pct_change = (REP16-REP14)/REP14*100, UNAF1416_pct_change = (UNAF16-UNAF14)/UNAF14*100,
    #2016-2018 pct_change
    DEM1618_pct_change = (DEM18-DEM16)/DEM16*100, REP1618_pct_change = (REP18-REP16)/REP16*100, UNAF1618_pct_change = (UNAF18-UNAF16)/UNAF16*100,
    #2018-2020 pct_change
    DEM1820_pct_change = (DEM20-DEM18)/DEM18*100, REP1820_pct_change = (REP20-REP18)/REP18*100, UNAF1820_pct_change = (UNAF20-UNAF18)/UNAF18*100,
    #2020-2022 pct_change
    DEM2022_pct_change = (DEM22-DEM20)/DEM20*100, REP2022_pct_change = (REP22-REP20)/REP20*100, UNAF2022_pct_change = (UNAF22-UNAF20)/UNAF20*100,
    #2022-2024 pct_change
    DEM2224_pct_change = (DEM24-DEM22)/DEM22*100, REP2224_pct_change = (REP24-REP22)/REP22*100, UNAF2224_pct_change = (UNAF24-UNAF22)/UNAF22*100,
    #2004-2024 pct_change
    DEM0424_pct_change = (DEM24-DEM04)/DEM04*100, REP0424_pct_change = (REP24-REP04)/REP04*100, UNAF0424_pct_change = (UNAF24-UNAF04)/UNAF04*100)
  
cleaner_compare_all <- compare_all |> 
  select(COUNTY, DEM0406_pct_change, REP0406_pct_change, UNAF0406_pct_change, DEM0608_pct_change, REP0608_pct_change, UNAF0608_pct_change, DEM0810_pct_change, REP0810_pct_change, UNAF0810_pct_change, DEM1012_pct_change, REP1012_pct_change, UNAF1012_pct_change, DEM1214_pct_change, REP1214_pct_change, UNAF1214_pct_change, DEM1416_pct_change, REP1416_pct_change, UNAF1416_pct_change, DEM1618_pct_change, REP1618_pct_change, UNAF1618_pct_change, DEM1820_pct_change, REP1820_pct_change, UNAF1820_pct_change,  DEM2022_pct_change, REP2022_pct_change, UNAF2022_pct_change, DEM2224_pct_change, REP2224_pct_change, UNAF2224_pct_change, DEM0424_pct_change, REP0424_pct_change, UNAF0424_pct_change)
```

```{r}
compare_all_pcts <-compare_all |>
  select(COUNTY,pct_DEM24,pct_REP24,pct_UNAF24,pct_DEM22,pct_REP22,pct_UNAF22,pct_DEM20,pct_REP20,pct_UNAF20, pct_DEM18,pct_REP18,pct_UNAF18,pct_DEM16,pct_REP16,pct_UNAF16,pct_DEM14,pct_REP14,pct_UNAF14,pct_DEM12,pct_REP12,pct_UNAF12,pct_DEM10,pct_REP10,pct_UNAF10,pct_DEM08,pct_REP08,pct_UNAF08,pct_DEM06,pct_REP06,pct_UNAF06,pct_DEM04,pct_REP04,pct_UNAF04)|>
  slice(25)

write_csv(compare_all_pcts, "october_data/compare_all_pcts.csv")

```

